Conversational expectations account for apparent limits on theory of mind use

### 2017-08-27 writing/journal_manuscript/pnas-submission.tex
Change bib name
Updated abstract

% TODO: Currently a bit too long (and probably not 'significant' sounding enough) 
\significancestatement{To communicate successfully, speakers must flexibly adjust the specificity of their utterances to the level demanded by the current context. In contexts where a speaker is unsure which objects their partner may be considering, it is helpful to provide additional disambiguating information: for example, saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands. 
We argue that influential arguments for limits on Theory of Mind (ToM) in adults have neglected the nuanced pragmatics of the communication tasks they rest upon. A careful analysis of these pragmatics demands using a rational model of communication suggests a minimal pair of replications showing that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}

@@

\begin{abstract}
Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life? Recent theories of pragmatic language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication. To reconcile these points of view, we focus attention on the problem faced by the \emph{speaker.} Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation. When we allow participants to naturally interact, we find that listener expectations are justified and errors are reduced. This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior--that is, to theory of mind. We formalize this argument in a computational model of pragmatic reasoning, and show that it provides a good fit with several recent findings.
\end{abstract}

@@

Humans can accurately and intelligently reason about the mental states of other humans. Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}. Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
%
While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking. In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably.

@@

@@

% Bibliography
\bibliography{pnas-sample}

>>>>

% TODO: Currently a bit too long (and probably too in the weeds...) 
\significancestatement{To communicate successfully, speakers must flexibly adapt %the specificity of 
their utterances to the demands of the current context. In contexts  which other objects are in view of their partnerrying to refer to an object in  , it is helpful to provide additional disambiguating information: for example, saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands. 
We argue that influential arguments for limits on Theory of Mind (ToM) in adults have neglected the nuanced pragmatics of the communication tasks they rest upon. A careful analysis of these pragmatic demands using a rational model of communication suggests a minimal pair of replications which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}

@@

\begin{abstract}
Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life? Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication. To reconcile these points of view, we use a computational model of communication under uncertainty to analyze the pragmatic demands of a common theory of mind test.  Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation, thus violating the listener's rational social expectations. This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind. 
\end{abstract}

@@

Humans can accurately and intelligently reason about the mental states of other humans. Theory of mind \cite{PremackWoodruff78_ChimpanzeeToM} allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}. 
%Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
%
While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking. In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably.

@@

%% Mention Westra & Carruthers as another critique focusing on pragmatics?

@@

% Bibliography
\bibliography{pnas-submission}
